---
title: "SMU 6306-Doing Data Science Case Study 1(Spring 2021)"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    theme: spacelab
    toc: yes
    toc_folat: yes
---



# Project Description

The purpose of this project is to study a dataset of 2410 beers produced by 558 
breweries in the United States.  We have data on the style, ABV (alcohol by volume)
and IBU (international bitterness units) of each beer, and we have the city
and state for each brewery.  The goal is to answer some specific questions 
about the data and to look for relationships or inferences which may be of
interest to a sponsor, Budweiser.

# Load Packages {.tabset}

These are all the packages that are used in the project to analyze the data and perform the agenda that were asked by the client.

## tidyverse
```{r}
#check if tidyverse is already installed or not and if it is, then load it.
if(!require(tidyverse))
  #If it is not installed. then tell R to install it.
  install.packages("tidyverse", dependencies = TRUE)
#Once it is installed, tell R to load it.
library(tidyverse)
```
## maps
```{r}
#check if maps is already installed or not and if it is, then load it.
if(!require(maps))
  #If it is not installed. then tell R to install it.
  install.packages("maps", dependencies = TRUE)
#Once it is installed, tell R to load it.
library(maps)
```
## caret
```{r}
#check if caret is already installed or not and if it is, then load it.
if(!require(caret))
  #If it is not installed. then tell R to install it.
  install.packages("caret", dependencies = TRUE)
#Once it is installed, tell R to load it.
library(caret)
```
## class
```{r}
#check if class is already installed or not and if it is, then load it.
if(!require(class))
  #If it is not installed. then tell R to install it.
  install.packages("class", dependencies = TRUE)
#Once it is installed, tell R to load it.
library(class)
```

# Study design and data processing

## Collection of the raw data
Data was provided by the client in two text .csv files.
1) Beers.csv
2) Breweries.csv

## Notes on the original (raw) data 
Beers.csv
2410 Observations.  
Name: Name of the beer.   
Beer_ID: Unique identifier of the beer.  
ABV: Alcohol by volume of the beer.  
IBU: International Bitterness Units of the beer.  
Brewery_ID: Brewery id associated with the beer.  
Style: Style of the beer.  
Ounces: Ounces of beer.  

Breweries.csv:  
558 Observations.  
Brew_ID: Unique identifier of the brewery.  
Name: Name of the brewery.  
City: City where the brewery is located.  
State: U.S. State where the brewery is located.  

# Description of the variables in the merged dataset

## Beer_name  
Name of the beer.   

## Beer_ID
Integer id number to uniquely identify each beer.

## ABV
Alcohol by volume (expressed as decimal fraction).

## IBU
International Bitterness Units of the beer (an integer).

## Brew_ID
Integer id number to uniquely identify the brewery which produced the beer.

## Style
Character string stored as a factor with 100 distinct values.

## Ounces
Serving size in ounces, represented as integer.

## Brewery_name
Character string stored as a factor with 551 distinct values.

## City
Character string stored as a factor with 384 distinct values.

## State
Two Character string stored as a factor with 51 distinct values.

# For finding some extra useful insights

## Collection of external data.

A new csv named "Alcohol_consumption.csv" is used having alcohol consumption per brewery Count.

## Notes on the external data
Alcohol_consumption.csv
51 observations
Rank: The ranks are assigned to each state according to highest to lowest consumption
State: State Name
Gallons Overall: The values given are in Millions

# Description of the variables in the merged dataset

## Rank
Integer id number to uniquely identify the rank of each state.

## State
Character string containing the state name

## Overall Gallons
Integer variable that has values in million gallons
